Salmon straying description

### 2017-05-03 manuscript/msv1.tex
mtheta ddm runs

@@

\begin{align}
  N_i(t+1) &= \left((1-m)N_i(t) + m N_j(t) \right){\rm e}^{-Z} \\ \nonumber
  &+ \left((R_i[\mu_i(t)|\theta_i]+\tilde{P}) (1-m)N_i(t) + (R_i[\mu_j(t)|\theta_i]+\tilde{P}) m N_j(t)\right) {\rm e}^{-\beta ((1-m)N_i(t) + m N_j(t))},
\end{align}

@@

\begin{align}
  R_i[\mu_i(t)|\theta_i] &= \int_{-\infty}^\infty r_{\rm max}\exp\left\{\frac{(x_i(t)-\theta_i)^2}{2\tau^2}\right\} {\rm pr}(x_i(t)|\mu_i,\sigma^2) {\rm d}x_i(t) \\ \nonumber
  &= \frac{r_{\rm max} \tau  }{\sqrt{\sigma ^2+\tau ^2}}\exp\left\{-\frac{(\theta_i-\mu_i(t))^2}{2 \left(\sigma ^2+\tau ^2\right)}\right\}.
\end{align}

@@

\begin{equation}
  m(t) = m_0 - m_0\left(\frac{N_i(t)}{C+N_i(t)}\right)
\end{equation}

@@

%Linking m/m0 and thetadiff

@@

%The effects of trait evolution :: trait heritability, trait variance, and habitat heterogeneity
%% Bifurcation to alternative stable state occurs at higher stray rates with increased trait heritability
%% There is a peak in the portfolio effect directly before this bifurcation is crossed (early warning signal)
%% When trait heritability is high, there is a greater sensitivity of total biomass to increasing stray rates, though a LOWER sensitivity of the difference between alternative steady state values for increasing stray rates (and vice versa)
%% For low heritability (h = 0.01 - 0.2 or so), moderate and insensitive changes in total biomass, but large differences between the values of alternative steady states leads to a local maximum for PE at low-moderate stray rates
%% Increased habitat heterogeneity
  %%% 1) exaggerates lowering of mean biomass with increasing stray rates
  %%% 2) increases the likelihood of alternative stable sites & increases the differences in alternative stable state values with increasing stray rates
  %%% 3) decreases the PE (!!)

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{figs/fig_Density.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{figs/fig_HSDcomb.png}
\caption{}
\end{figure}

>>>>

%\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}

@@

\begin{align}
  N_i(t+1) &= \left((1-m)N_i(t) + m N_j(t) \right){\rm e}^{-Z} \\ \nonumber
  &+ \left((R_i[\mu_i(t)|\theta_i]+\tilde{P}) (1-m)N_i(t) + (R_i[\mu_j(t)|\theta_i]+\tilde{P}) m N_j(t)\right) {\rm e}^{-\beta ((1-m)N_i(t) + m N_j(t))},
  \label{eq:N}
\end{align}

@@

\begin{align}
  R_i[\mu_i(t)|\theta_i] &= \int_{-\infty}^\infty r_{\rm max}\exp\left\{\frac{(x_i(t)-\theta_i)^2}{2\tau^2}\right\} {\rm pr}(x_i(t)|\mu_i,\sigma^2) {\rm d}x_i(t) \\ \nonumber
  &= \frac{r_{\rm max} \tau  }{\sqrt{\sigma ^2+\tau ^2}}\exp\left\{-\frac{(\theta_i-\mu_i(t))^2}{2 \left(\sigma ^2+\tau ^2\right)}\right\}.
  \label{eq:R}
\end{align}

@@

\begin{equation}
  m(t) = m_0\left(1- \frac{N_i(t)}{C+N_i(t)}\right),
  \label{eq:ddm}
\end{equation}

@@

%Linking m/m0 and thetadiff
Stray rate is intrinsically linked to the distance between the local and straying population.
The greater the distance between two populations, the lower the expected rate of straying (REF).
We can account for this interdependence in our model by assuming that $m$ (if the stray rate is constant) or $m_0$ (if the stray rate is density dependent) is a function of $theta_i-theta_j$, which can be assumed to be large if the remote site $j$ is a great distance away from the local site $i$.
If sites $i$ and $j$ are very close, the stray rate is assumed to maximized at $m,m_0 = 0.5$.
Thus, we can integrate these two variables by setting $m,m_0 = (2 + \epsilon (\theta_i-\theta_j))^{-1}$, where $\epsilon$ sets the sensitivity of a declining $m$ to increasing distance (greater values of $\theta_i-\theta_j$).

@@

%The effects of trait evolution :: trait heritability, trait variance, and habitat heterogeneity
%% Bifurcation to alternative stable state occurs at higher stray rates with increased trait heritability
%% There is a peak in the portfolio effect directly before this bifurcation is crossed (early warning signal), and then a decline in PE with high m
%% When trait heritability is high, there is a greater sensitivity of total biomass to increasing stray rates, though a LOWER sensitivity of the difference between alternative steady state values for increasing stray rates (and vice versa)
%% For low heritability (h = 0.01 - 0.2 or so), moderate and insensitive changes in total biomass, but large differences between the values of alternative steady states leads to a local maximum for PE at low-moderate stray rates
%% Increased habitat heterogeneity
  %%% 1) exaggerates lowering of mean biomass with increasing stray rates
  %%% 2) increases the likelihood of alternative stable sites & increases the differences in alternative stable state values with increasing stray rates
  %%% 3) decreases the PE (!!)

%Alternative stable states
Straying generally lowers steady state densities for both local and remote populations. %, regardless of trait heritability and variance or habitat heterogeneity
The decline in steady state densities is not gradual: as straying increases, the system crosses a fold bifurcation whereby the single steady state splits and becomes two alternative steady states: one at high biomass density, and one at low biomass density (Fig. \ref{}).
This bifurcation occurs at lower values of the stray rate $m$ for lower levels of trait heritability $h^2$, indicating that tighter coupling between ecological and evolutionary dynamics  requires higher rates of straying to give rise to alternative stable states.

Trait heritability also has a large impact on the sensitivity of both the total steady state density ($N^*_T=N^*_1+N^*_2$) as well as the difference between steady state densities ($\Delta N=\sqrt{(N^*_1-N^*_2)^2}$).
Greater trait heritability (high $h^2$) results in a larger decline in $N_T^*$ for increasing stray rates $m$, though a moderate change in $\Delta N$ with increasing stray rates.
Conversely, if the trait has a low heritability, an increase in the stray rate has little impact on the total biomass density, though very large impacts on the difference in population densities between sites.

Together these differences in steady state population densities as a function of trait heritability and the rate of straying between populations results in nonlinear changes to the portfolio effect, quantified as

\begin{equation}
  {\rm PE} = \frac{\overline{{\rm CV}}_{i,j}}{{\rm CV}_T}.
  \label{eq:pe}
\end{equation}

\noindent We define $\overline{{\rm CV}}_{i,j}$ as the average coefficient of variation across the local population $i$ and the remote population $j$, and ${\rm CV}_T$ as the coefficient of variation of the total or aggregated population.
Higher ${\rm PE}$ corresponds to an increased potential for ecological rescue between populations, thus buffering the system as a whole from extinction, whereas the lowest possible value is ${\rm PE_min}=1$

In the region where there is a single steady state value there is a correspondingly high portfolio effect.
As the fold bifurcation is approached with greater rates of straying, the portfolio effect spikes due to a large increase in the standard deviation of individual trajectories.
This explosion in variance is a well-known phenomenon that occurs near a fold bifurcation and lies at the heart of early warning signal research (REFS).
Directly to the right of the fold bifurcation (in Fig \ref{fig:PE}C) where there are two alternative steady state densities, the portfolio effect is minimized if heritability is low, and it remains elevated if heritability is high.
As the stray rate increases to the right of the bifurcation, increased trait heritability results in a steady decrease in the portfolio effect.
If trait hertitability is low, the portfolio effect changes nonlinearly: after the spike associated with the fold bifurcation, {\rm PE} first increases at intermediate stray rates and then declines as $m\rightarrow0.5$.

% Density dependent m
%% Because m[t] @ steady state 0<m[t]<m0, we observe the same dynamics for higher m0 as we do for lower constant m.
%% Higher rates of straying are tolerable (there is a not a decline in PE for very high m0 as there is for constant m)

% When m and thetadiff are linked
%%Constant m
  %%% Low stray rate (high theta1-theta2) leads to alternative stable states and relatively low PE ~ sharp transition
  %%% High stray rate (low theta1-theta2) increases the steady state (and no A.S.S.)... 
  %%% This suggests that space naturally buffers against the negative effects of mixing dissimilar populations, however it is NOT SMOOTH
  %%% This is particularly true for low heritability. High heritability eliminates the A.S.S. region
%% Similar results for DDm

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{figs/fig_Density.pdf}
\caption{} \label{fig:traj}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{figs/fig_MDPE_hm.pdf}
\caption{} \label{fig:PE}
\end{figure}

\begin{figure}
\centering
\begin{subfigure}[t]{0.8\textwidth}
\centering
\includegraphics[width=\textwidth]{figs/fig_MDPE_hm_theta3.pdf} 
\caption{Low habitat heterogeneity ($\Delta\theta=3$)} \label{fig:hetero1}
\end{subfigure}

\begin{subfigure}[t]{0.8\textwidth}
\centering
\includegraphics[width=\textwidth]{figs/fig_MDPE_hm_theta8.pdf} 
\caption{High habitat heterogeneity ($\Delta\theta=8$)} \label{fig:hetero2}
\end{subfigure}

 \caption{Changes in total means, difference in means, and the portfolio effect for different habitat heterogeneities}

\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{figs/fig_MDPE_hm_cross358.pdf}
\caption{} \label{fig:PE}
\end{figure}
